---
title: "MTcars regression project"
author: "c zingler"
date: "16/06/2020"
output:
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Motor Trends Cars Linear Regression Analysis  
## Executive Summary  
We will be analyzing the Motor Trend Car Road Tests data, it  comprises 
fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).  
This data was compiled in 1974.
Simply put we’re asking the question, which types of cars: manual or
automatic transmission cars goes the furthest, using (mpg) as an
indication.
The findings are:
Manual cars are more fuel efficient.  For this Data set they go on average 7.245 miles further per gallon.


## Project requirement
We are to approach this task as a typical business initiated investigation,
thus we can not assume a through understanding of statistic by the audience

### Data description

Data Table name: mtcars  

Format
A data frame with 32 observations on 11 (numeric) variables.

[, 1]	mpg         Miles/(US) gallon  
[, 2]	cyl	        Number of cylinders  
[, 3]	disp        Displacement (cu.in.)  
[, 4]	hp          Gross horsepower  
[, 5]	drat        Rear axle ratio  
[, 6]	wt	        Weight (1000 lbs)  
[, 7]	qsec        1/4 mile time  
[, 8]	vs	        Engine (0 = V-shaped, 1 = straight)  
[, 9]	am	        Transmission (0 = automatic, 1 = manual)  
[,10]	gear        Number of forward gears
[,11]	carb        Number of carburetors  

### Preparing the data for analysis 

```{r ehco=TRUE}
##load the graphics pakage in case of need
library(ggplot2)
## Load the data of interest
data(mtcars)
## Now we need to ensure some of the columns are factors for easier linear model fitting
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)

## show the data
head(mtcars,10)

```
## Exploritory analysis  

From the Scatter plot we can see the is a strong relationship between MPG 
and Cyl, Disp, Hp, Drat, Wt and am.  This needs some quantification, please
read below for subsequent discussion.
In particular the relationship between MPG and Automatic and Manual
transmissions.

So lets look at MPG under a single regressor of interest.
```{r echo=TRUE,strip.white=TRUE}
basefit <- lm(mpg~am,data = mtcars)
summary(basefit)
##and
basefit1 <- lm(mpg~am-1,data = mtcars)
summary(basefit1)

```

So there is a 7.245 decrease in MPG for Automatic transmissions, and the 
second linear model details the MPG mean for each type of transmission.
This can be verified by inspecting Appendix Fig.2

Now lets investigate a linear model of factors of interest.
But note disp and wt are related, as well as drat and hp being related. So run the model without them.

```{r echo=TRUE, strip.white=TRUE}
bestfit <- lm(mpg~cyl+hp+wt+am,data = mtcars)
summary(bestfit)
## and now for there means
bestfit1 <- lm(mpg~cyl+hp+wt+am-1,data = mtcars)
summary(bestfit1)
```

So all components are quite significant as the model has an R-squared of
86.59%. A very high degree of correlation

Now lets perform an ANOVA to compare the models

```{r echo=TRUE, strip.white=TRUE}
anova(basefit,bestfit)

```

So the P-value is very small and quite significant, therefore
We reject the Null hypothesis that WT, cyl, and hp do not contribute to the model accuracy.

### Inference on effect of Transmition on MPG
lets perform a T Test.
```{r echo=TRUE,strip.white=TRUE}
t.test(mpg~am,data = mtcars)

```

And we see that the Type of transmission has a significant impact on the
level of Miles pre Gallon achieved. that is greater than a 95% confidence,
as seen in the T Test.

## Appendix 
### Residual, Diagnostics and Plots

We compute some regression diagnostics on the bestfit model residuals

```{r echo=TRUE,strip.white=TRUE, fig.width=6.5, fig.height=5.5}
par(mfrow = c(2,2))
plot(bestfit)
```

From the above plots, we can make the following observations,

* The points in the Residuals vs. Fitted plot seem to be randomly scattered
on the plot and verify the independence condition. 
* The Normal Q-Q plot consists of the points which mostly fall on the line
indicating that the residuals are normally distributed. 
* The Scale-Location plot consists of points scattered in a constant band
pattern, indicating constant variance.
* There are some distinct points of interest (outliers or leverage points)
in the top right of the plots. 

We now compute some regression diagnostics of our model to find out these
interesting leverage points as shown in the following section.  
```{r echo=TRUE,strip.white=TRUE}
Lev <- hatvalues(bestfit)
tail(sort(Lev),3)
resinf <- dfbetas(bestfit)
tail(sort(resinf[,6]),3)
```
The same cars are mentioned in the residual plots, our analysis looks
correct.    
```{r echo=FALSE, strip.white=FALSE, fig.width=6.5, fig.height=3.9}
par(mfcol =c(1,1))
plot(mtcars)

```
Scatter plots matrix of MTCARS dataset  
```{r echo=FALSE ,strip.white=TRUE ,fig.width=6.5, fig.height=2.5}
par(mfcol =c(1,2))
boxplot(mpg~am ,data = mtcars, col=(c("pink","cyan")), xlab = "Transmisson  ( 0= Automatic, 1= Manual", ylab = "Miles per Gallon", main= "Fig 2 MPG vs Transmission")
boxplot(mpg~cyl, data = mtcars, col=(c("red","green","blue")), xlab= "Number of Cylinders", ylab= "miles per gallon", main= "Fig 3 MPG Vs Cylinders")
```
e
